Relative parabolic factorizations of Chevalley groups

### 2015-04-05 stability-conditions.tex
Typesetting, fixes and examples for Stability conditions section

\subsection{Stable rank}
Recall the definition of {\it stable rank} introduced by H.~Bass and L.~Vaserstein (see~\cite{Ba64, Va69}).
\begin{dfn} Recall that a row $a=(a_1,\ldots, a_n)\in {}^n\!R$ is called {\it $I$-unimodular} if elements $a_1-1, a_2, \ldots, a_n$ are contained in $I$ while $a_1, a_2, \ldots, a_n$ generate $R$ as an ideal.\end{dfn}
A column $b \in R^n$ will be called $I$-unimodular if its transpose $b^t$ is an $I$- unimodular row. We denote the set of all $I$-unimodular rows (resp. columns) by $\Umd(n,R,I)$ (resp. $\Ums(n,R,I)$).
When $R=I$ we refer to $R$-unimodular rows and columns as simply unimodular.

\begin{lemma} \label{lemma:relstrlemma} For any $I$-unimodular row $a=(a_1, \ldots, a_n)$ there exists $I$-unimodular column $b=(b_1,\ldots, b_n)^T$ such that $ab = \sum\limits_{i=1}^n a_i b_i = 1$. \end{lemma}

\begin {dfn} An $I$-unimodular row $a=(a_1,\ldots a_{n+1})$ is called {\it stable} if one can choose $b_1,\ldots, b_n\in R$ such that
row $(a_1 + a_{n+1}b_1,\ldots, a_{n}+ a_{n+1}b_n)$ is also $I$-unimodular. if, moreover, elements $b_1,\ldots, b_n$ can be chosen from  $I$ then such $a$ will be called {\it $I$-stable}. \end{dfn}

We say that a pair $(R, I)$ satisfies stable range condition $\SR_n(R, I)$ if any $I$-unimodular row of length $n+1$ is stable.
\begin{lemma}\label{lemma:relstrlemma2}\strut\begin{enumerate}\item The condition $\SR_n(R, I)$ implies $\SR_m(R,I)$ for any $m\geq n$.
\item The condition $\SR_n(R, I)$ implies that any $I$-unimodular row of length $n+1$ is $I$-stable.\end{enumerate}\end{lemma}

By definition, the {\it relative stable rank} $\sr(R, I)$ of a pair $(R, I)$ is the smallest natural number $n$ such that condition $\SR_n(R, I)$ holds.
If $\SR_n(R, I)$ is not satisfied for any $n>0$ we set $\sr(R, I)=\infty$.

The following proposition summarizes basic properties of stable ranks.
\begin{prop} \label{prop:sr_properties} Let $R$ be arbitary commutative unital ring and let $I\trianglelefteq R$ be its ideal.
 \begin{itemize}
  \item For any ideal $J\trianglelefteq R$, $J\subseteq I$ one has $$\sr(R, I)\leq\sr(R),\quad \sr(R/J, I/J)\leq \sr(R, I).$$
  \item Stable rank of a direct product of rings is equal to the maximum of stable ranks of its factors: $$\sr(\prod\limits_{i=1}^n R_i) = \max\limits_{i=1}^n\left(\sr(R_i)\right).$$
  \item Stable rank does not change after taking quotient modulo nil-radical: $\sr(R)=\sr(R/\Rad(R))$.
 \end{itemize}\end{prop} \begin{proof} See~\cite{Va71}. \end{proof}

\begin{example} Since the stable rank of a field is one, one can conclude from the previous proposition that $\sr(R)=1$ for any semilocal ring $R$. \end{example}

\subsection{Absolute stable rank}

%TODO: $\asr(R,I)$

>>>>

\subsection{Stable rank}
Recall the definition of {\it stable rank} introduced by H.~Bass and L.~Vaserstein (see~\cite{Ba64, Va69}).
\begin{dfn}
Recall that a row $a=(a_1,\ldots,a_n)\in{}^n\!R$ is called {\it $I$-unimodular} if elements $a_1-1,a_2,\ldots,a_n$ are contained in $I$ while $a_1,a_2,\ldots,a_n$ generate $R$ as an ideal.
\end{dfn}
A column $b \in R^n$ will be called $I$-unimodular if its transpose $b^t$ is an $I$-unimodular row. We denote the set of all $I$-unimodular rows (resp. columns) by $\Umd(n,R,I)$ (resp. $\Ums(n,R,I)$).
When $R=I$ we refer to $R$-unimodular rows and columns as unimodular.

\begin{lemma}\label{lemma:relstrlemma}
For any $I$-unimodular row $a=(a_1,\ldots,a_n)$ there exists $I$-unimodular column $b=(b_1,\ldots,b_n)^t$ such that $ab=\sum\limits_{i=1}^n a_i b_i=1$.
\end{lemma}

\begin {dfn} An $I$-unimodular row $a=(a_1,\ldots a_{n+1})$ is called {\it stable} if one can choose $b_1,\ldots,b_n\in R$ such that the row $(a_1+a_{n+1}b_1,\ldots,a_n+a_{n+1}b_n)$ is also $I$-unimodular. If, moreover, elements $b_1,\ldots,b_n$ can be chosen from  $I$ then such $a$ will be called {\it $I$-stable}. \end{dfn}

We say that a pair $(R,I)$ satisfies stable range condition $\SR_n(R,I)$ if any $I$-unimodular row of length $n+1$ is stable.
\begin{lemma}\label{lemma:relstrlemma2}\strut
\begin{enumerate}
\item The condition $\SR_n(R,I)$ implies $\SR_m(R,I)$ for any $m\geqslant n$.
\item The condition $\SR_n(R,I)$ implies that any $I$-unimodular row of length $n+1$ is $I$-stable.
\end{enumerate}
\end{lemma}

By definition, the {\it relative stable rank} $\sr(R,I)$ of a pair $(R,I)$ is the smallest natural number $n$ such that condition $\SR_n(R,I)$ holds.
If no such $n$ exists, we set $\sr(R,I)=\infty$.

The following proposition summarizes the basic properties of stable ranks.
\begin{prop} \label{prop:sr_properties}
Let $R$ be arbitrary commutative unital ring and let $I\trianglelefteq R$ be its ideal.
\begin{itemize}
\item For any ideal $J\trianglelefteq R$, $J\subseteq I$ one has 
\[ \sr(R,I)\leqslant\sr(R),\quad\sr(R/J,I/J)\leqslant\sr(R,I). \]
\item Stable rank of a direct product of rings is equal to the maximum of stable ranks of its factors:
\[ \sr\left(\prod\limits_{i=1}^n R_i\right)=\max\limits_{i=1}^n\left(\sr(R_i)\right). \]
\item Stable rank does not change after taking quotient modulo the nil-radical: $\sr(R)=\sr(R/\Rad(R))$.
\end{itemize}
\end{prop}
\begin{proof} See~\cite{Va71}. \end{proof}

\begin{example}
Since the stable rank of a field is one, one can conclude from the previous proposition that $\sr(R)=1$ for any semilocal ring $R$.
\end{example}
\begin{example}
Other examples of rings of stable rank $1$ include boolean rings, the ring of algebraic integers, the ring of entire functions, the disk-algebra.
\end{example}
\begin{example}
If $J$ is the Jacobson radical of $R$, one has $\sr(R,J)=1$.
\end{example}

\subsection{Absolute stable rank}
For a row $a=(a_1,\ldots,a_n)\in{}^n\!R$ denote by $\mathfrak{I}(u)$ the intersection of ideals of $R$ containing $a_1,\ldots,a_n$.
It is easy to see that a row $a\in R^n$ is unimodular if and only if $\mathfrak{I}(u)=R$. 
Clearly, for any $g\in\GL(n,R)$ one has $\mathfrak{I}(u\cdot g)=\mathfrak{I}(u)$.

A ring $R$ is said to satisfy condition $\ASR_{n}(R,I)$ if for any row $a$ such that $a_1-1,\ldots,a_{n+1}\in I$ one can find elements $b_1,\ldots b_n\in R$ such that
\[ \mathfrak{I}(a_1+b_1a_{n+1},\ldots,a_n+b_na_{n+1})^t = \mathfrak{I}(a_1,\ldots,a_{n+1})^t. \]
It is easy to see that $\ASR_m(R,I)$ implies $\ASR_n(R,I)$ for any $n\geqslant m$. 
By definition, the {\it absolute relative stable rank} $\asr(R,I)$ is the smallest natural $n$ such that $\ASR_n(R, I)$ holds. Similarly, we set $\asr(R,I)=\infty$ if no such $n$ exists. When $I=R$ we call $\asr(R, R)$ simply the {\it absolute stable rank} of $R$ and denote it by $\asr(R)$.

Let $R$ be a commutative ring. We denote by $\Max(R)$ its {\it maximum spectrum} i.e. the set of maximal ideals of $R$ equipped with Zariski topology.
The following result gives an upper estimate for the absolute stable rank.~\cite[Theorem~2.3]{EO}, \cite[Theorem~3.7]{MKV}).
\begin{thm}
Let $R$ be a commutative ring whose maximum spectrum $\Max(R)$ can be covered with a finite family of subsets $X_1\ldots, X_m$ such that the topological dimension of every subset $X_i$ does not exceed some number $d$. Then $\asr(A)\leqslant d+1$.
\end{thm}
Clearly one has
\[ \sr(R,I)\leqslant\asr(R,I)\leqslant\asr(R)\leqslant \dim(\Max(R))+1\leqslant\dim(\Spec(R))+1. \]
